Legion: Expressing Locality and Independence with Logical Regions

### 2012-07-02 doc/popl2013/langdef.tex
checkpoint

\newcommand{\oton}[1]{{#1}_1,\ldots,{#1}_n}
\newcommand{\otok}[2]{{#2}_1,\ldots,{#2}_{#1}}
\newcommand{\dplus}{\text{++}}

@@

$T$ & ::= &  & types \\
  &$\mid$& bool $\;\;\;\mid\;\;\;$ int & base types \\
  &$\mid$& $\langle T_1, T_2 \rangle$ & tuple \\
  &$\mid$& $T@(\oton{r})$ & pointer \\
  &$\mid$& $\text{coloring}(r)$ & region coloring \\
  &$\mid$& $\exists \oton{r}. T\text{ where }\Omega$ & region relationship \\
  &$\mid$& $\forall \oton{r}. (\oton{T}), \Phi \rightarrow T_r$ & functions \\
\\
$\Omega$ & ::= & $\{ \oton{\omega} \}$ & region constraints \\
$\omega$ & ::= & $r_1 \leq r_2$ & subregion \\
  &$\mid$& $r_1 * r_2$ & disjointness \\
\\
$\Phi$ & ::= & $\{ \oton{\phi} \}$ & privileges \\
$\phi$ & ::= & reads$(r)$ $\;\;\;\mid\;\;\;$ writes$(r)$ $\;\;\;\mid\;\;\;$ reduces$(id,r)$ & \\
\\
$v$ & ::= & & values \\
  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & base values \\
  &$\mid$& $\langle v_1, v_2 \rangle$ & tuple \\
  &$\mid$& null $\;\;\;\mid\;\;\;$ $a$ & address \\
  &$\mid$& $\{ (a_i, iv), \ldots \}$ & coloring \\
  &$\mid$& $\langle \langle \oton{\rho}, v\rangle \rangle$ & region relationship instance \\
\\
$bv$ & ::= & false $\;\;\;\mid\;\;\;$ true \\
\\
$iv$ & ::= & 0 $\;\;\;\mid\;\;\;$ 1 $\ldots$ \\
\\
$e$ & ::= & & expressions \\
  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & constants \\
  &$\mid$& $\langle e_1, e_2 \rangle$ $\;\;\;\mid\;\;\;$ $e$.1 $\;\;\;\mid\;\;\;$ $e$.2 & tuple \\
  &$\mid$& $id$ &  \\
  &$\mid$& $\text{new}\ T@r$ $\;\;\;\mid\;\;\;$ $\text{null }T@r$ $\;\;\;\mid\;\;\;$ $\text{isnull}(e)$ & \\
  &$\mid$& $\text{upregion}(e, r_1,\ldots,r_n)$ $\;\;\;\mid\;\;\;$ $\text{downregion}(e, r_1,\ldots,r_n)$ & \\
  &$\mid$& $\text{read}(e_1)$ $\;\;\;\mid\;\;\;$ $\text{write}(e_1, e_2)$ $\;\;\;\mid\;\;\;$ $\text{reduce}(id, e_1, e_2)$ & memory access \\
  &$\mid$& $\text{newcolor}\ r$ $\;\;\;\mid\;\;\;$ $\text{color}(e_1, e_2, e_3)$ & coloring operations \\
  &$\mid$& $e_1 + e_2$ & integer operations \\
  &$\mid$& $e_1 < e_2$ & comparison operations \\
  &$\mid$& $\text{let}\ id : T = e_1 \text{in}\ e_2$ &  \\
  &$\mid$& $\text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3$ &  \\
  &$\mid$& $id[r_1, \ldots, r_n](e_1,\ldots,e_n)$ & task calls \\
  &$\mid$& $\text{partition}\ r_p\text{ using }e_1\text{ as }\oton{r}\text{ in }\ e_2$ &  \\
  &$\mid$& $\text{pack}\ e_1\ \text{as}\ T[r_1,\ldots,r_n]$ &  \\
  &$\mid$& $\text{unpack}\ e_1\ \text{as}\ id : T[r_1,\ldots,r_n]\ \text{in}\ e_2$ &  \\

@@

\finfrule{
\begin{array}{l}
\Gamma(id) = \forall \otok{k}{r'}.(\oton{T}),\Phi_i \rightarrow T_r \\
\typeenvx e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\
\Phi_i[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*
\end{array}}
{\typeenvx id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]} \\

@@

where: 
\begin{itemize}
\item $M$ is the mapping from logical regions $r_i$ in types to physical regions $\rho_i$, which are sets of memory locations
\item $L$ is the mapping of local variable names to their values
\item $H$ is the heap typing, mapping from memory locations to types
\item $S$ is the store, mapping from memory locations $l_i$ to values
\item $C$ is the ``clobber set'', a set of memory locations that may be accessed by concurrently executing operations (??? does this need to be split into read/write/reduce?)
\item $e$ is the expression being evaluated
\item $v$ is the result of the evaluation
\item $E$ is a memory trace (i.e. the effects) of the execution, a list of read/write/reduce operations in the order they were performed
\end{itemize}

@@

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = apply(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \langle e_1, e_2 \rangle \mapsto \langle v_1, v_2 \rangle, E_1 \dplus E_2} \\

@@

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E \\
S' = \text{apply}(S, E, C) \\
v = S'[l]
\end{array}}
{\opsenvx \text{read}(e_1) \mapsto v, E \dplus [ read(l, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenvx \text{write}(e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ write(l, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenvx \text{reduce}(id, e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ reduce_{id}(l, v) ]} \\

@@

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto l, E_2 \\
S'' = \text{apply}(S', E_2, C) \\
\opsenvx[S=S''] e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \}
\end{array}}
{\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E_1 \dplus E_2 \dplus E_3} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = v_1 + v_2
\end{array}}
{\opsenvx e_1 + e_2 \mapsto v', E_1 \dplus E_2} \hspace{1cm}

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = \begin{cases}
true, & \text{if $v_1 < v_2$}. \\
false, & \text{otherwise}.
\end{cases}
\end{array}}
{\opsenvx e_1 < e_2 \mapsto v', E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto true, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto false, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_3 \mapsto v_3, E_3
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E_1 \dplus E_3} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S_1 = \text{apply}(S, E_1, C) \\
\opsenvx e_2 \mapsto v_2, E_2 \\
\ldots \\
S_n = \text{apply}(S_{n-1}, E_n, C) \\
task id[\otok{k}{r'}](f_1 : T_1, \ldots, f_n : T_n) : T_r, E = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \\
L' = \{ (f_1, v_1), \ldots, (f_n, v_n) \} \\
\opsenvx[M=M',L=L',S=S_n] e_{n+1} \mapsto v_{n+1}, E_{n+1}
\end{array}}
{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E_1 \dplus \ldots \dplus E_{n+1}} \\

\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
\rho_i = \{ l : (l, i) \in K \} \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[M=M',S=S'] e_2 \mapsto v, E_2
\end{array}}
{\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E_1 \dplus E_2} \\

@@

\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E_1 \dplus E_2 } \\

@@

If
\begin{enumerate}
\item $\typeenvx e : T$
\item $\opsenvx e \mapsto v, E$
\end{enumerate}
we would like to prove that all of the dynamic effects described by $E$ are
included in the static privileges $\Phi$.  A judgement $E : \Phi$ means that
for all effects $\epsilon$ in $E$,
\begin{enumerate}
\item if $\epsilon = read(l, v)$, then for some $r$, $reads(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = write(l, v)$, then for some $r$, $writes(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = reduce(id, l, v)$, then for some $r$, $reduces(id,r) \in \Phi \wedge l \in M(r)$
\end{enumerate}

@@

\newcommand{\llbracket}{[\![}
\newcommand{\rrbracket}{]\!]}

The mapping $M$ is a function from logical region names $r$ to physical regions $\rho$.  It will
be convenient to apply $M$ to types and even functions that operate on types by defining
$\hat x = M \llbracket x \rrbracket$ to be $x$ with all logical region names $r_i$ replaced by their
corresponding physical region $\rho_i = M(r_i)$.  Note that $M \llbracket x \rrbracket$ is not
well-formed if it contains a logical region name $r_j \not\in dom(M)$. \\

@@

@@

>>>>

\newcommand{\oton}[1]{{#1}_1,\ldots,{#1}_n}
\newcommand{\otok}[2]{{#2}_1,\ldots,{#2}_{#1}}
\newcommand{\dplus}{\text{++}}
\newcommand{\llbracket}{[\![}
\newcommand{\rrbracket}{]\!]}

@@

$T$ & ::= &  & types \\
  &$\mid$& bool $\;\;\;\mid\;\;\;$ int & base types \\
  &$\mid$& $\langle T_1, T_2 \rangle$ & tuple \\
  &$\mid$& $T@(\oton{r})$ & pointer \\
  &$\mid$& $\text{coloring}(r)$ & region coloring \\
  &$\mid$& $\exists \oton{r}. T\text{ where }\Omega$ & region relationship \\
  &$\mid$& $\forall \oton{r}. (\oton{T}), \Phi, Q \rightarrow T_r$ & functions \\
\\
$\Omega$ & ::= & $\{ \oton{\omega} \}$ & region constraints \\
$\omega$ & ::= & $r_1 \leq r_2$ & subregion \\
  &$\mid$& $r_1 * r_2$ & disjointness \\
\\
$\Phi$ & ::= & $\{ \oton{\phi} \}$ & privileges \\
$\phi$ & ::= & reads$(r)$ $\;\;\;\mid\;\;\;$ writes$(r)$ $\;\;\;\mid\;\;\;$ reduces$(id,r)$ & \\
\\
$Q$ & ::= & $\{ \oton{q} \}$ & coherence modes \\
$q$ & ::= & atomic$(r)$ $\;\;\;\mid\;\;\;$ simult$(r)$ & \\
\\
$v$ & ::= & & values \\
  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & base values \\
  &$\mid$& $\langle v_1, v_2 \rangle$ & tuple \\
  &$\mid$& null $\;\;\;\mid\;\;\;$ $a$ & address \\
  &$\mid$& $\{ (a_i, iv), \ldots \}$ & coloring \\
  &$\mid$& $\langle \langle \oton{\rho}, v\rangle \rangle$ & region relationship instance \\
\\
$bv$ & ::= & false $\;\;\;\mid\;\;\;$ true \\
\\
$iv$ & ::= & 0 $\;\;\;\mid\;\;\;$ 1 $\ldots$ \\
\\
$e$ & ::= & & expressions \\
  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & constants \\
  &$\mid$& $\langle e_1, e_2 \rangle$ $\;\;\;\mid\;\;\;$ $e$.1 $\;\;\;\mid\;\;\;$ $e$.2 & tuple \\
  &$\mid$& $id$ &  \\
  &$\mid$& $\text{new}\ T@r$ $\;\;\;\mid\;\;\;$ $\text{null }T@r$ $\;\;\;\mid\;\;\;$ $\text{isnull}(e)$ & \\
  &$\mid$& $\text{upregion}(e, r_1,\ldots,r_n)$ $\;\;\;\mid\;\;\;$ $\text{downregion}(e, r_1,\ldots,r_n)$ & \\
  &$\mid$& $\text{read}(e_1)$ $\;\;\;\mid\;\;\;$ $\text{write}(e_1, e_2)$ $\;\;\;\mid\;\;\;$ $\text{reduce}(id, e_1, e_2)$ & memory access \\
  &$\mid$& $\text{newcolor}\ r$ $\;\;\;\mid\;\;\;$ $\text{color}(e_1, e_2, e_3)$ & coloring operations \\
  &$\mid$& $e_1 + e_2$ & integer operations \\
  &$\mid$& $e_1 < e_2$ & comparison operations \\
  &$\mid$& $\text{let}\ id : T = e_1 \text{in}\ e_2$ &  \\
  &$\mid$& $\text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3$ &  \\
  &$\mid$& $id[r_1, \ldots, r_n](e_1,\ldots,e_n)$ & task calls \\
  &$\mid$& $\text{partition}\ r_p\text{ using }e_1\text{ as }\oton{r}\text{ in }\ e_2$ &  \\
  &$\mid$& $\text{pack}\ e_1\ \text{as}\ T[r_1,\ldots,r_n]$ &  \\
  &$\mid$& $\text{unpack}\ e_1\ \text{as}\ id : T[r_1,\ldots,r_n]\ \text{in}\ e_2$ &  \\

@@

\finfrule{
\begin{array}{l}
\Gamma(id) = \forall \otok{k}{r'}.(\oton{T}),\Phi_i, Q \rightarrow T_r \\
\typeenvx e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\
\Phi_i[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*
\end{array}}
{\typeenvx id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]} \\

@@

where: 
\begin{itemize}
\item $M$ is the mapping from logical regions $r_i$ in types to physical regions $\rho_i$, which are sets of memory locations.  It will be useful to apply this logical-to-physical mapping to other
things that defined in terms of logical regions to yield a corresponding definition in terms of
physical regions.  The notation $\hat x = M \llbracket x \rrbracket$ is used to describe the
result of replacing all free logical region names $r_i$ in $x$ with the corresponding physical
region $\rho_i$.  (Note that $M \llbracket x \rrbracket$ is not
well-formed if it contains a free logical region name $r_j \not\in dom(M)$.)
\item $L$ is the mapping of local variable names to their values
\item $H$ is the heap typing, mapping from memory locations to types
\item $S$ is the store, mapping from memory locations $l_i$ to values
\item $C$ is the ``clobber set'', a set of memory locations that may be accessed by concurrently executing operations (??? does this need to be split into read/write/reduce?)
\item $e$ is the expression being evaluated
\item $v$ is the result of the evaluation
\item $E$ is a memory trace (i.e. the effects) of the execution, a list of read/write/reduce operations in the order they were performed.  An operation $\epsilon$ is one of the following:
\begin{itemize}
\item $read(l, c, v)$, a read of location $l$ with coherence mode $c$, returning value $v$
\item $write(l, c, v)$, a write to location $l$ with coherence mode $c$, storing value $v$
\item $reduce_{id}(l, c, v)$, a reduction into location $l$ with coherence mode $c$ using a
commutative function $id$ with value $v$
\end{itemize}
where coherence modes are one of ({\em excl}, {\em atomic}, {\em simult}).

The following helper functions are used in some of the rules that deal with memory effects:
\begin{itemize}
\item $apply(S, E, C)$ is used to determine the effect on the store of a sequence of memory operations
\item $mark\_coherence(E, \hat Q)$ marks the memory operations in a subtask's trace to
constrain how the operations in that trace can be interleaved with other sibling tasks' operations
\item $valid\_interleave(S, C, E', \oton{E})$ is used to determine if $E'$ is a valid way to
interleave the memory traces of two or more subexpressions
\end{itemize}
Figure~\ref{fig:opsemfns} gives the definitions of {\em apply} and {\em mark\_coherence}.  For
the {\em valid\_interleave} predicate, we state some of the necessary conditions here, but leave 
the construction of actual valid interleavings until a later section.  Informally, a interleaving
of two or more memory traces must include all the operations from each trace (and no additional
operations), with each trace's operations remaining in order.  A valid interleaving must further
satisfy the requirement that the value read from any location that's not in the clobber set $C$
yield the most recently stored/reduced value.

\begin{figure*}
\centering{
$
\begin{array}{lll}
apply(S, [], C) & = & S \\
apply(S, E \dplus [ read(l, c, v) ]) & = & apply(S, E, C) \\
apply(S, E \dplus [ write(l, c, v) ]) & = & apply(S, E, C)[v/l] \\
apply(S, E \dplus [ reduce_{id}(l, c, v) ]) & = & apply(S, E, C)[id(S[l], v)/l] \\
mark\_coherence([], \hat Q) & = & [] \\
mark\_coherence(E \dplus [ op(l, c, v) ], \hat Q) & = & mark\_coherence(E, \hat Q) \dplus \\
& & op(l, c', v), \text{ where } \\
& & c' =
\begin{cases}
simult, & \text{if }\exists \rho. l \in \rho \wedge simult(\rho) \in \hat Q \\
atomic, & \text{if }\exists \rho. l \in \rho \wedge atomic(\rho) \in \hat Q \\
excl, & \text{otherwise}
\end{cases} \\
any\_interleave(E', [], \ldots, []) & \rightarrow & true \\
any\_interleave([ op(l, c, v) ] \dplus E', E_1, \ldots, [ op(l, c, v) ] \dplus E_i, \ldots, E_n) & \rightarrow & any\_interleave(E', E_1, \ldots, E_i, \ldots, E_n) \\
valid\_interleave(S, C, E', \oton{E}) & \rightarrow & any\_interleave(E', \oton{E}) \\
valid\_interleave(S, C, E' \dplus [ read(l, c, v) ] \dplus E'', \oton{E}) & \rightarrow & l \in C \vee S'(l) = v, \text{ where } S' = apply(S, E', C)
\end{array}
$}
\caption{Operational Semantics Helper Functions}
\label{fig:opsemfns}
\end{figure*}

A few helper functions are used in the operational semantics rules, and defined in 
A number of rules will also use the property $valid\_interleave(S, C, E', E_1, E_2)$ to 
constrain how the memory operations of two subexpressions can be interleaved.  We will 
consider actual constructions of valid interleavings later.  For now, we simply rely on
the following necessary conditions for any valid interleaving:
\begin{enumerate}
\item A valid interleaving $E'$ of $E_1$ and $E_2$ contains exactly all operations from both
$E_1$ and $E_2$, in some order.
\item The operations in each of $E_1$ and $E_2$ remain in order in $E'$.
\item For every $read(l, c, v)$ in $E'$ for which $l$ is not in the clobber set $C$, the value 
returned by the read is consistent with the original store $S$ and any preceding writes or
reductions.
\end{enumerate}
\end{itemize}

@@

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = apply(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \langle e_1, e_2 \rangle \mapsto \langle v_1, v_2 \rangle, E'} \\

@@

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E \\
S' = \text{apply}(S, E, C) \\
v = S'[l]
\end{array}}
{\opsenvx \text{read}(e_1) \mapsto v, E \dplus [ read(l, excl, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenvx \text{write}(e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ write(l, excl, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenvx \text{reduce}(id, e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ reduce_{id}(l, excl, v) ]} \\

@@

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto l, E_2 \\
S'' = \text{apply}(S', E_2, C) \\
\opsenvx[S=S''] e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \} \\
valid\_interleave(S, C, E', E_1, E_2, E_3)
\end{array}}
{\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E'} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = v_1 + v_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx e_1 + e_2 \mapsto v', E'} \hspace{1cm}

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = \begin{cases}
true, & \text{if $v_1 < v_2$}. \\
false, & \text{otherwise}.
\end{cases} \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx e_1 < e_2 \mapsto v', E'} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E'} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto true, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E'} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto false, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_3 \mapsto v_3, E_3 \\
valid\_interleave(S, C, E', E_1, E_3)
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E'} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S_1 = \text{apply}(S, E_1, C) \\
\opsenvx e_2 \mapsto v_2, E_2 \\
\ldots \\
S_n = \text{apply}(S_{n-1}, E_n, C) \\
valid\_interleave(S, C, E', \oton{E}) \\
S' = \text{apply}(S, E', C) \\
task id[\otok{k}{r'}](f_1 : T_1, \ldots, f_n : T_n), \Phi_i, Q : T_r = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \\
L' = \{ (f_1, v_1), \ldots, (f_n, v_n) \} \\
\opsenvx[M=M',L=L',S=S'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \\
E'_{n+1} = mark\_coherence(E_{n+1}, M \llbracket Q \rrbracket) \\
valid\_interleave(S, C, E'', E', E_{n+1})
\end{array}}
{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''} \\

\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
\rho_i = \{ l : (l, i) \in K \} \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[M=M',S=S'] e_2 \mapsto v, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E'} \\

@@

\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E' } \\

@@

An important property of the legion type system is that any expression that is well-typed is
guaranteed to access the heap only in ways permitted by the privileges under which it was typed.
A judgement $M \vdash E : \Phi$
If
\begin{enumerate}
\item $\typeenvx e : T$
\item $\opsenvx e \mapsto v, E$
\end{enumerate}
we would like to prove that all of the dynamic effects described by $E$ are
included in the static privileges $\Phi$.  A judgement $E : \Phi$ means that
for all effects $\epsilon$ in $E$,
\begin{enumerate}
\item if $\epsilon = read(l, v)$, then for some $r$, $reads(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = write(l, v)$, then for some $r$, $writes(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = reduce(id, l, v)$, then for some $r$, $reduces(id,r) \in \Phi \wedge l \in M(r)$
\end{enumerate}

@@

The mapping $M$ is a function from logical region names $r$ to physical regions $\rho$.  It will
be convenient to apply $M$ to types and even functions that operate on types by defining
$\hat x = M \llbracket x \rrbracket$ to be $x$ with all logical region names $r_i$ replaced by their
corresponding physical region $\rho_i = M(r_i)$.   \\

@@

\finfrule
{\begin{array}{l}
E_1 \sim H \hspace{1cm}
E_2 \sim H \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{ E' \sim H } \\

@@

The soundness proof above depended, in part, on the properties of a valid interleaving of two
memory traces.  We now consider algorithms that create interleavings that satisfy those
conditions.

The simplest way to achieve a valid interleaving is to concatenate the traces in order.  This
will always result in a valid interleaving, but implies a serialization of the evaluation of
subexpressions.

\framebox{$\forall S, C, E_1, E_2. valid\_interleave(S, C, E_1 \dplus E_2, E_1, E_2)$}

Next step: For noninterfering traces $E_1 \# E_2$, show any interleaving is valid.

Third step: For interactions only on atomic-coherence locations, serializability is sufficient.
